% Acknowledgements

\pdfbookmark[1]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer

\begin{flushright}{\slshape    
You're never gonna grow if you don't grow now \\
You're never gonna know if you don't find out \\
You're never going back never turning around \\
You're never gonna go if you don't go now \\
}
\end{flushright}

\bigskip

%----------------------------------------------------------------------------------------

\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}

\noindent

It goes without saying that my deepest gratitude goes to my Ph.D. advisor Pierre Geurts, without whom this thesis would not exist. I can't thank you enough for all the help you have given me throughout my doctoral journey; academically speaking, as well as on a personal level. Thank you for putting your trust in me and welcoming me in your research group, for always being excited about my research, for always finding the time to meet me, and most importantly for granting me the research freedom I needed to pull this Ph.D. off. You already know how grateful I am of having had you as my supervisor and always on my side during my Ph.D. Thank you also to Mike Kestemont, for not only being part of my jury, but for all the hard work you have put in the INSIGHT project, without which this manuscript would be at least twice as short; thank you for always giving me valuable feedback about my research, and for all our conversations that have allowed me to keep my Vlaams sharp. An immense thank you goes to the other jury members, which I have been lucky enough to have as part of my thesis committee. Thank you Nicu Sebe, for being the first academic person spiking my interest in machine learning back then when I was a BSc. student in Italy. Thank you Aske Plaat, for being so enthousiastic about my work and for the giving my lots of reseach ideas that I would like to explore in the future, I will try to get a policy gradient version of DQV working as soon as possible. Thank you Raphael Fonteneau, for the extremely pleasant scientific discussion we had at my defense and for always being nice to me each time I encountered you in Montefiore's corridors, your smile always put me in a good mood. Thank you Gilles, for all our coffees, the gossipping, the nice times spent in Canada, and for always being there for me if I needed someone to complain to, or confuse with some random reinforcement learning idea. Lastly, thank you Louis, for being the president of my defense, and for all the stimulating scientific discussions we had throughout the years.     

My non-academic life in Liege was made easier by some very special people which I was fortunate enough to encounter and share time with: first and foremost I wish to thank my Italian landlord Domenico, who has helped me tremendeosly when I first moved from Groningen to Liege. Thank you for helping me out furnishing my apartment, for the Friday night poker evenings at your bar, and last but not least, for renting me your apartment at such a nice price. A special thank you also goes to my Indian neighbours Ravi, Sridevi and their wonderful family, owners of the delicious "Indian Flavour" restaurant: thank you for all the amazing vegetarian food that you have served me throughout the years, and for always being there when I needed to have a chat in English. I also wish to thank my swimming teacher Constance: having a swim in the coast of Bretagne with all those fishes is one of those memories which will always have a special place in my heart. I consider learning how to swim the second largest achievement of the time spent in Liege, so thank you for all your patience and for all those coupons which I never found the time to use.  

A big thank you goes to those friends who have always been there for me despite the physical distance separating us: Francesco, thank you for all of our calls and for constantly showing me your support even throughout my darkest days; I finally promise I will be visiting you in Tokyo as soon as possible. I still think it should be possible to train a Generative Adversarial Network on only a small sample of the input distribution, so get your coding skills back together. Irene, thank you for always being there for a call from either Piombinodese or London and thank you for your patience and kindness. Samuele, thank you for being always ready to comment the newest season of Masterchef with me, and for always being down for a solid scientific discussion, whether it was about machine learning or women. A special thanks also goes to my favourite Bayesian friend Matteo aka "Bijoux", now that you have finally learned how to code I'm sure you have a great scientific future ahead of you, and I can't wait to read your future papers as a Ph.D. student. Remember that until the day of your defense I will always be one step ahead of you in the academic foodchain. I also owe my gratitude to my dear Dutch friend Marco Gunnink, I still remember as if it was yesterday our first road trip from Groningen to Liege right before the start of my Ph.D. journey, and our super-sketchy experiences in the streets of Liege. Also thank you for always making me feel welcome at your place in Groningen when I was feeling home-sick, for visiting me in Liege at New Year's and for still lending me money when I'm having issues with my bank account. Lastly thank you to all my Groningen friends, older and newer ones, who to this day make me feel home: thank you to my favourite optimal-control scientist Kat, who is the goto person if you want to know what the acronyms of a control-theory conference stay for, and thank you to Laurids, for our always fruitful discussions, about machine learning, food, and of course, chess; I really outplayed your Dutch defense right before my defense, didn't I?!       

A special merçi goes to the entire Montefiore research unit for heartly welcoming me back in 2017 when I first joined the department. A special thank you goes to Pierre's research group; Romain, Van Anh, Antonio, Jean Michel, Yann; thank you for your company around the lunch table and for always doing your best in steering the conversation towards English instead of French. I would also like to thank you for the countless scientific discussions we had, without which, most of the contributions within this thesis would have never seen the light. I would also like to thank Antoine, for being a good friend when I was having a hard time, the amazing times spent in Montreal and Vancouver when visiting Canada for NeurIPS, and for the nice road trip from Liege to Groningen. Michael, thank you for being my office neighbour, for always being there for a nice chat and for the amazing pictures you took at my defense. Remy; thank you for the (long) jogging sessions in Tilff and for our discussions ranging from object detection to soccer. Arthur, thank you for all the saturday afternoon's you've spent at my place (or should you be thanking me?), for your always inspiring curiosity and (mostly annoying) questions about reinforcement learning. Thank you also to Maxime, Malavika, Navdip, Raphael, Ulysse, Laurine, Pascal and Nicolas, for simply being around Montefiore and always being down for a nice chat.

Thank you to two very special persons which helped me immensively over the last stages of my Ph.D. Thank you Thomas, for the help you gave me every Friday afternoon by taking care of my mental health issues when I needed it the most. Thank you Federica, for your everlasting willpower, courage, and the inspiring passion you put in everything you do. Now that a new exciting chapter of your life is about to start I'm sure a fulfilling journey is awaiting you and I'm as curious as you are to see where this will lead you.

Last, but not least, thank you to my special family: thank you mum for the strong, independent woman you are and for the constant support you have shown me throughout my long academic journey which has taken me away from Italy. Thanks to my amazing grandparents which have always offered me shelter when I needed to switch from the Walloon-Belgian experience to the Dutch-Belgian experience. Thank you grandpa for being so patient with my grandma constantly forgetting what Ph.D. stays for and for never giving up on reminding her. Un ultimo grazie to my cousin Pino and to my aunt Lina, who are still convinced I'm the kind of physician Dr. and who keep asking me for medical advice I'm not able to give. 

 


\endgroup
